Effectiveness of private sector malaria control at a large sugar

### 2017-10-20 inst/rmd/maragra_preliminary_analysis.Rmd
big update

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
bairros <- maragra::bairros
census <- maragra::census
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
mc <- maragra::mc
workers <- maragra::workers
```

@@

For the purposes of this first pass, we "bin" IRS exposure into 5 groups: never, ever but 180+ days ago, 90-80 days, 60-90 days, and in the last 60 days.

@@

```{r}
# Create model data
model_data <-
  ab_panel %>%
  filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  left_join(matched) %>%
  mutate(days_since = ifelse(is.na(days_since), 'Never',
                             ifelse(days_since > 180, '180+',
                                    ifelse(days_since > 90, '090-180',
                                           ifelse(days_since > 60, '060-090',
                                                  '<060'))))) %>%
  mutate(quarter = lendable::quarter_extract(date)) %>%
  mutate(quarter = as.character(quarter)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
# Need to add seasonality
# Need to adjust for houses on and off site
fit <- glm(absent ~ days_since + quarter,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit))
ors <- exp(confint(fit))
ors <- data.frame(ors)
names(ors) <- c('Lower', 'Upper')

@@

## To do

- More robust models and checks.  
- Limit ton only clinical malaria. 
- Un-bin IRS (ie, estimate semi-parametric curve of effectiveness waning)  
- More investigation of worker type, shift, etc.

>>>>

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
bairros <- maragra::bairros
census <- maragra::census
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
irs <- maragra::irs
mc <- maragra::mc
workers <- maragra::workers
```

@@

For the purposes of this first pass, we "bin" IRS exposure into 5 groups: before IRS (includes IRS > 365 days ago), 180+ days ago, 90-80 days, 60-90 days, and in the last 60 days.

@@

```{r}
# Create model data
model_data <-
  ab_panel %>%
  filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  left_join(matched) %>%
  mutate(days_since = ifelse(days_since > 180, '180+',
                                    ifelse(days_since > 90, '090-180',
                                           ifelse(days_since > 60, '060-090',
                                                  ifelse(days_since >= 0, '000-060',
                                                         ifelse(days_since < 0, 'Before', NA)))))) %>%
  filter(!is.na(days_since)) %>%
  mutate(days_since = factor(days_since, levels = c('Before', '000-060', '060-090', '090-180', '180+'))) %>%
  mutate(quarter = lendable::quarter_extract(date)) %>%
  mutate(quarter = as.character(quarter)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
# Need to add seasonality
# Need to adjust for houses on and off site
fit <- glm(absent ~ days_since + quarter,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit))
ors <- exp(confint(fit))
ors <- data.frame(ors)
names(ors) <- c('Lower', 'Upper')

@@

## Discontinuity analysis (not using propoensity score matching)

```{r}
# Create model data for all obs (not just propensity score)
model_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  # left_join(matched) %>%
  mutate(days_since = ifelse(days_since > 180, '180+',
                                    ifelse(days_since > 90, '090-180',
                                           ifelse(days_since > 60, '060-090',
                                                  ifelse(days_since >= 0, '000-060',
                                                         ifelse(days_since < 0, 'Before', NA)))))) %>%
  filter(!is.na(days_since)) %>%
  mutate(days_since = factor(days_since, levels = c('Before', '000-060', '060-090', '090-180', '180+'))) %>%
  mutate(quarter = lendable::quarter_extract(date)) %>%
  mutate(quarter = as.character(quarter)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) #%>%
  # # remove the period 30 days after spraying
  # filter(days_since >= 30 | days_since <= 0)
# Need to add seasonality
# Need to adjust for houses on and off site
```

#### Raw absenteeism as a function of days since IRS

```{r}
x <- model_data %>%
  group_by(days_since) %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100)

ggplot(data = x,
       aes(x = days_since,
           y = absenteeism_rate)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkblue') +
  theme_maragra() +
  labs(x = 'Days since IRS',
       y = 'Absenteeism rate') +
  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
```

#### Raw absenteeism as a function of days since IRS by quarter

```{r}
x <- model_data %>%
  group_by(days_since, quarter) %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100)

ggplot(data = x %>% mutate(quarter = paste0('Q', quarter)),
       aes(x = days_since,
           y = absenteeism_rate)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkblue') +
  theme_maragra() +
  labs(x = 'Days since IRS',
       y = 'Absenteeism rate') +
  # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) +
  facet_wrap(~quarter) +
  theme(axis.text.x = element_text(angle = 90))
```

#### Modeling without propensity score

(Absence)

```{r}
# Need to adjust for houses on and off site
fit <- glm(absent ~ days_since + quarter,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit))
ors <- exp(confint(fit))
ors <- data.frame(ors)
names(ors) <- c('Lower', 'Upper')

(Sick absence)

#### Visualization of absenteeism before / after IRS

```{r}
plot_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  mutate(quarter = lendable::quarter_extract(date)) %>%
  mutate(quarter = as.character(quarter)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x <- plot_data %>%
  group_by(months_since) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)

ggplot(data = x,
       aes(x = months_since,
           y = absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'Before/after IRS',
       subtitle = 'All absences') +
  theme_maragra()

ggplot(data = x,
       aes(x = months_since,
           y = sick_absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'Before/after IRS',
       subtitle = 'Sick absences only') +
  theme_maragra()
```

```{r}
plot_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  mutate(quarter = lendable::quarter_extract(date)) %>%
  mutate(quarter = as.character(quarter)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x <- plot_data %>%
  group_by(months_since, quarter) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)

ggplot(data = x %>% mutate(quarter = paste0('Q', quarter)),
       aes(x = months_since,
           y = absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  facet_wrap(~quarter) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate') +
  theme_maragra()
```

### 2017-10-20 inst/rmd/paper.Rmd
big update

---
title: Effectiveness of private sector malaria control at a large sugar
  facility in Southern Mozambique
csl: elsevier-harvard.csl
date: ''
output:
  html_document:
    css: resources/style.css
    pandoc_args:
    - +RTS
    - -K16000m
    - -RTS
    toc: yes
  pdf_document:
    toc: yes
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

@@

<!-- <hr> -->
<!-- <p style="text-align:left;"> -->
<!-- [http://economicsofmalaria.com](http://economicsofmalaria.com) -->
<!-- <span style="float:right;">joebrew@gmail.com</span> -->
<!-- </p> -->
<!-- <hr> -->

@@

@@

```{r}
# # Create model data
# model_data <-
#   ab_panel %>%
#   left_join(irs) %>%
#   left_join(workers %>%
#               dplyr::select(oracle_number,
#                             sex,
#                             race_code,
#                             department,
#                             permanent_or_temporary)) %>%
#   mutate(days_since = ifelse(is.na(days_since), 'Never',
#                              ifelse(days_since > 180, '180+',
#                                     ifelse(days_since > 90, '090-180',
#                                            ifelse(days_since > 60, '060-090', 
#                                                   '<060')))))
# # Need to add seasonality
# # Need to adjust for houses on and off site
# fit <- glm(absent ~ days_since,
#            family = binomial('logit'),
#           data = model_data)
# exp(coef(fit))
```

@@

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown'
), 'packages.bib')
```

>>>>

---
title: Effectiveness of private sector malaria control at a large sugar
  facility in Southern Mozambique
date: ''
output: 
  pdf_document:
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

@@

@@

```{r}
authors <- data.frame(name = c('Joe Brew',
                               'Kizito Gondo',
                               'Elton Dorkin',
                               'Eduardo Nhamahanga',
                               'Menno Pradhan',
                               'Laia Cirera',
                               'Ranjeeta Thomas',
                               'Elisa Sicuri'),
                      email = c('joe@databrew.cc',
                                'KGondo@illovo.co.za',
                                'EDorkin@illovo.co.za',
                                'ENhamahanga@illovo.co.za',
                                'm.p.pradhan@vu.nl',
                                'laia.cirera@isglobal.org',
                                'ranjeeta.thomas@imperial.ac.uk',
                                'elisa.sicuri@isglobal.org'),
                      affilitation = c("isglobal,cism,vu",
                                       'ma',
                                       'ma',
                                       'ma',
                                       "vu,uva",
                                       "icl",
                                       'isglobal,cism',
                                       "isglobal,icl,cism"),
                      footnote = c("Corresponding Author",
                                   rep('', 7)),
                      stringsAsFactors = FALSE)
institutions <- data.frame(code = c('isglobal',
                                    'icl',
                                    'cism',
                                    'vu',
                                    'uva',
                                    'ma'),
                           institution = c('Barcelona Institute for Global Health',
                                           'Imperial College London',
                                           'Centro de Investigação em Saúde de Manhiça',
                                           'VU University Amsterdam',
                                           'University of Amsterdam',
                                           'Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd'),
                           address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
                                       'South Kensington Campus, London SW7 2AZ, U.K.',
                                       'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
                                       'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
                                       'REC E, Roetersstraat 11, Amsterdam, Netherlands',
                                       'CP 2789, Maputo, Mozambique'),
                      stringsAsFactors = FALSE)

make_author <- function(authors,
                        institutions){

  out <- c()
  for (i in 1:nrow(authors)){
    this_author <- authors$name[i]
    these_institutions <- authors$affilitation[i]
    these_institutions <- unlist(strsplit(these_institutions, ','))
    author_text <- paste0(this_author)
    for (j in 1:length(these_institutions)){
      this_institution <- these_institutions[j]
      if(any(grepl(this_institution, out))){
        author_text <- paste0(author_text,
                              '\\footrecall{',
                              this_institution,
                              '}'
                              )
      } else {
        # new institution, get full name
        full_name <- institutions$institution[institutions$code == this_institution]
        address <- institutions$address[institutions$code == this_institution]
        author_text <- paste0(author_text,
                              '\\footremember{',
                              this_institution,
                              '}{',
                              full_name,
                              ': ',
                              address,
                              '}')
      }
    }
    out[i] <- author_text
  }
  out <- paste0(out, collapse = ', ')
  cat(out)
}
```

\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_author(authors = authors,
            institutions = institutions)
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

Here goes some abstract text. Bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla .

\end{changemargin}
\end{center}

@@

### Identification strategy

For our purposes we are analyzing the effect of one intervention (IRS) on 2 outcomes (absence and illness) with many confounders (age, worker type, seasonality, etc.). Our analysis can be visualized formulaically as follows:

$$
\begin{equation}
\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
\end{equation}
$$

Our outcome is probabilistic and binomial (ie, one is either absent or present / infected or not infected). Our demographic confounders (represented by $...$) will be a function of iterative model selection. Our intervention (IRS) is not a simple yes/no, but rather the product of whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment (represented above as the interaction term, where where $_t$ represents time elapsed since commencement of the most recent IRS campaign).

```{r}
# Create model data
model_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  # left_join(matched) %>%
  mutate(days_since = ifelse(is.na(days_since), 'Never',
                             ifelse(days_since > 180, '180+',
                                    ifelse(days_since > 90, '090-180',
                                           ifelse(days_since > 60, '060-090',
                                                  '<060'))))) %>%
  mutate(quarter = lendable::quarter_extract(date)) %>%
  mutate(quarter = as.character(quarter)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))
# Need to add seasonality
# Need to adjust for houses on and off site
fit <- glm(absent ~ days_since + quarter,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit))
ors <- exp(confint(fit))
ors <- data.frame(ors)
names(ors) <- c('Lower', 'Upper')

ors <- cbind(x, ors)
ors$Variable <- row.names(ors)
ors <- ors %>%
  dplyr::rename(OR = x) %>%
  dplyr::select(Variable, OR, Lower, Upper)
# Combine into a df
kable(ors) %>%
  kableExtra::kable_styling(full_width = FALSE)
```

We run the same model, but instead of estimating absences, we estimate only the likelihood of sick absences. The results (in form of odds ratios) are below.

```{r}
# Need to add seasonality
# Need to adjust for houses on and off site
fit <- glm(absent_sick ~ days_since + quarter,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit))
ors <- exp(confint(fit))
ors <- data.frame(ors)
names(ors) <- c('Lower', 'Upper')

ors <- cbind(x, ors)
ors$Variable <- row.names(ors)
ors <- ors %>%
  dplyr::rename(OR = x) %>%
  dplyr::select(Variable, OR, Lower, Upper)
# Combine into a df
kable(ors) %>%
  kableExtra::kable_styling(full_width = FALSE)

```

@@

```{r, eval=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown'
), 'packages.bib')
```
